---
title: "Covid19 project"
author: "Deepanshu Bhasin"
date: "07/04/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 8. APPENDIX (RCode)


```{r warning=FALSE,message=FALSE}
#getwd()
library(ggplot2)
library(janitor)
library(dplyr)
library(tidyverse)
library(caret)
library(corrplot)
library(DMwR)
library(reshape2)
library(e1071)
library(MLmetrics)
library(scatterplot3d)
library(randomForest)
library(mlbench)
```

##SECTION 0
Importing data
```{r}
data <- read.csv('F:/Deepanshu/WESTERNU/SUBJECTS/Winter2020/Stats Mod 2/project/dataset.csv')
```
Replacing unnecessary spaces and other characters that can further become problematic
```{r}
data <- data %>% 
  rename_all(., ~janitor::make_clean_names(.))
#str(data)
```


## SECTION 1 EXPLORATORY DATA ANALYSIS
### 1.1 Data Cleaning

Figure 1.1a - Creating visualization report to check proportion of missing values
```{r warning=FALSE,message=FALSE, include=TRUE, fig.height=15, fig.width=12}
na_cols = apply(is.na(data),2,sum)
na_dfs = data.frame("key" = colnames(data), "count_NA" = na_cols)
na_dfs = na_dfs[with(na_dfs, order(-count_NA)),]
ggplot(na_dfs, aes(reorder(key,count_NA),count_NA/5644,fill = count_NA)) +
  geom_bar(stat = "identity") + theme_minimal() + coord_flip()+
  labs(title = "", x = "Variables",y='% of Missing Values', fill= "Count")
```

It can be noted that there is a strong presence of missing data in many attributes.

Removing 100% empty columns
```{r}
total_NA_cols <- data %>%
  select_if(~sum(is.na(.x))== nrow(data))%>%
  colnames()

total_NA_cols
data <- data %>%
  dplyr::select(-one_of(total_NA_cols))
```

## TASK1
Removing unnecessary columns like Patient_ID and target variables for Task2 from data 
```{r}
data_reduced<-data %>%
  dplyr::select(-c(patient_id,patient_addmited_to_regular_ward_1_yes_0_no,patient_addmited_to_semi_intensive_unit_1_yes_0_no,patient_addmited_to_intensive_care_unit_1_yes_0_no))
```

Removing 100% empty rows-NO EXAM TAKEN
```{r}
total_NA_rows  <- data_reduced %>%
  select_if (~!sum(!is.na(.x))==nrow(data_reduced))%>%
  {apply (.,1 ,function(x){sum(is.na(x))})!= ncol(.)}%>%
  which ()

data_reduced <-  data_reduced %>% slice(total_NA_rows)
```

Converting all categories to character
```{r}
data_reduced  <- 
  data_reduced %>%
  mutate_if(is.factor,as.character)
```

Missing values in character variables
```{r}
data_reduced%>%
  select_if(is.character) %>% 
  purrr::map(~table(.,useNA ='ifany'))
```

```{r}
count_along_row_NA <- which((apply(is.na(data_reduced),1,sum) == 63))
data_reduced <- data_reduced[-count_along_row_NA, ]

dim(data_reduced)
```

Removing response for time being
```{r}
exam_res<-data_reduced$sars_cov_2_exam_result
data_reduced<-data_reduced[,-2]
```

## Transformation
Converting to numeric
```{r}
data_reduced$urine_leukocytes <- as.numeric(ifelse(data_reduced$urine_leukocytes == "<1000", 999, data_reduced$urine_leukocytes));

#patient_age_quantile
data_reduced$patient_age_quantile<-as.numeric(data_reduced$patient_age_quantile)
```

Separating categorical and numerical data
```{r}
type_var<-sapply(data_reduced, typeof)

index_col <- which(type_var=="character"|type_var=="logical")

data_categ <-data_reduced[,index_col]

data_numeric<-data_reduced[,-index_col]
```

categorical character data
Converting categorical to factors
```{r}
data_categ<-data.frame(lapply(data_categ,as.factor))
#summary(data_categ)
```

```{r}
levels_categ <- lapply(data_categ,levels)

#Transform levels into numeric
data_categ<-data.frame(lapply(data_categ,as.numeric))
```

```{r}
#produced all numeric data
data_all_num<-cbind(data_categ,data_numeric)
```
```{r}
#added response variable back to dataset and renamed it as 'task1_resp'
data_task1<-cbind(exam_res,data_all_num)
data_task1<-data_task1%>%
  rename(task1_resp=exam_res)
```

Figure 1.1b - Checking count of missing values
```{r, include=TRUE, fig.height=12, fig.width= 12}
apply(is.na(data_task1), 2, sum) %>% 
  sort() %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(variable = rowname, na_count = V1) %>% 
  ggplot(aes(x = reorder(variable, na_count), y = na_count)) +
  geom_col() + 
  geom_text(aes(x = variable, y = na_count + 35, label = na_count, angle = 90), size = 3.5) +
  labs(x = 'variable', title = 'Variables ordered by NA count') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

Selected first few columns and removed columns having count of NAs greater than 7
```{r}
min_NA_col<-apply(is.na(data_task1),2, sum)

data_task1<-data_task1[,min_NA_col<=7]

```

Dropping rows still having NA values in columns
```{r}
data_task1<-data_task1%>%
  drop_na()
```

### 1.2 Visualizations
Fugure 1.2a - Checking correlation
```{r include=TRUE, fig.height=22, fig.width= 40}
corrplot::corrplot(cor(data_task1[,-1]),method="number",title = "Correlation plot",
                   type = "upper",number.cex = .8, tl.cex = 1.5)
```

Removing medical test data with high correlation
```{r}
data_task1<-data_task1[-c(2:20)]
#colnames(data_task1)
```

Removing clinical urine test data with high correlation
```{r}
data_task1<-data_task1[-c(3:8)]
#colnames(data_task1)
```
```{r}
data_task1<-data_task1[-c(4:10)]
#colnames(data_task1)
```

Figure 1.2b - checking multicollinearity for few variables to keep only 1 of them
```{r include=TRUE, fig.height=12, fig.width= 12}
par(mfrow=c(1,4))
plot(data_task1$mean_corpuscular_hemoglobin_concentration_mch,data_task1$mean_corpuscular_volume_mcv)
plot(data_task1$hemoglobin,data_task1$red_blood_cells)
plot(data_task1$hematocrit,data_task1$red_blood_cells)
plot(data_task1$hematocrit,data_task1$hemoglobin)
```

Removing columns to avoid multicollinearity
```{r}
data_task1<-data_task1%>%
  dplyr::select(-c(hematocrit, hemoglobin,
            mean_corpuscular_hemoglobin_concentration_mchc,
            mean_corpuscular_hemoglobin_mch, mean_corpuscular_volume_mcv))
```

Fig 1.2c Checking Data Balance for response
```{r}
names(data_task1$task1_resp)<-NULL
freq_res<-table(data_task1$task1_resp)
pie(freq_res/sum(freq_res))

#table to show number of +ve & -ve cases
table(data_task1$task1_resp)
```

Figure 1.2d - Distribution of positive vs negative 
```{r}
pos_case <- subset(data_task1, data_task1$task1_resp == "positive")
neg_case <- subset(data_task1, data_task1$task1_resp == "negative")

# positive cases by age
ggplot (pos_case, aes(x = as.factor(patient_age_quantile), fill = 1)) +
  geom_bar (stat = "count", position = position_dodge(), show.legend = FALSE) +
  theme_classic () + labs (x = "Age class", y = "Number") + 
  labs(title = 'Positive cases by Age class')

# negative cases by age
ggplot (neg_case, aes (x = as.factor(patient_age_quantile), fill = 1)) +
  geom_bar (stat = "count", position = position_dodge (), show.legend = FALSE) +
  labs(title = 'Negative cases by Age class')+
  theme_classic () + labs (x = "Age class", y = "Number")
```

Figure 1.2e - Boxplot to handle outliers
```{r include=TRUE, fig.height=18, fig.width= 30}
data_task1 %>% 
  pivot_longer(-c(1,4), names_to = 'variable', values_to = 'value') %>% 
  ggplot(aes(x = variable, y = value, fill = task1_resp)) +
  stat_boxplot(geom = 'errorbar', size = 0.3) +
  geom_boxplot() +
  labs(title = 'Boxplot of covariates by group') + 
  theme(axis.text.x = element_text(size = 20,angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```
handling outliers
```{r}
neg_case%>%
   dplyr::select(basophils,platelets,eosinophils,red_blood_cell_distribution_width_rdw)%>%
  summary()
```

removing observations directly
```{r}
outliers_neg <- c(which(data_task1$platelets > 9),
         which(data_task1$basophils > 4),
         which(data_task1$eosinophils > 5),
         which(data_task1$red_blood_cell_distribution_width_rdw > 6))
outliers_neg

data_task1<-data_task1%>%
  slice(-outliers_neg)
```


##TASK2
### PART 2 OF EDA  

Continuing from section 1.1 after removing 100% null columns
```{r}
data2<-data%>%
  filter(patient_addmited_to_intensive_care_unit_1_yes_0_no!=0|patient_addmited_to_regular_ward_1_yes_0_no!=0|patient_addmited_to_semi_intensive_unit_1_yes_0_no!=0)
dim(data2)
```

Checking missing values for dataset
```{r, include=TRUE, fig.height=12, fig.width= 20}
apply(is.na(data2), 2, sum) %>% 
  sort() %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(variable = rowname, na_count = V1) %>% 
  ggplot(aes(x = reorder(variable, na_count), y = na_count)) +
  geom_col() + 
  geom_text(aes(x = variable, y = na_count + 35, label = na_count, angle = 90), size = 3.5) +
  labs(x = 'variable', title = 'Variables ordered by NA count') +
  theme(axis.text.x = element_text(size=15,angle = 90, hjust = 1, vjust = 0.5))
```

As we have seen already the correlation between categorical variables and few of the numeric variables we will again remove those variables from our data and dropping rows with NA values for those columns
```{r}
data2<-data2 %>%
  dplyr::select('patient_addmited_to_regular_ward_1_yes_0_no',
                'patient_addmited_to_semi_intensive_unit_1_yes_0_no',
                'patient_addmited_to_intensive_care_unit_1_yes_0_no',
                'patient_age_quantile','basophils','eosinophils',
                'leukocytes','lymphocytes','monocytes','platelets',
                'red_blood_cell_distribution_width_rdw','red_blood_cells') %>% drop_na()

#colnames(data2)
```

## FEATURE ENGINEERING
```{r}
data2<-data2%>%
  rename(regular=1,semi=2,intensive=3)
```

We will bind the patient admitted to different wards into single variable "admitted_to" which will be the response variable for this task and will have factors as regular,semi-intensive,intensive
```{r}
data2 <- data2 %>% 
  pivot_longer(c(regular,semi,intensive),names_to="admitted_to") %>% 
  filter(value!= 0) %>% 
  dplyr::select(admitted_to, everything()) %>% 
  mutate(admitted_to = factor(admitted_to,levels =c('regular','semi','intensive'))) %>% 
  dplyr::select(-value)
```

Visualizations
Checking proportions of different levels
```{r}
table(data2$admitted_to)
```

Figure 1.2f
```{r out.width = '90%', fig.width = 13, fig.asp = 1.2}
data2%>%pivot_longer(-1, names_to = 'variable', values_to = 'value') %>%
  mutate(variable = as.factor(variable)) %>% 
  ggplot(aes(x = variable, y = value)) +
  geom_violin(aes(fill = admitted_to)) +
  facet_wrap(~variable, scales = 'free', ncol = 3) +
  scale_fill_manual(values = c('#00AFBB', '#E7B800', '#FC4E07')) +
  labs(title = 'Distribution of features by wards') +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(size = 16))
```
We can see outliers are present but this time will not remove it as already we have less data


Figure 1.2g -  Scatterplot showing distribution of wards in data for few variables
```{r warning=FALSE,message=FALSE}
colors <- c('red', 'blue', 'green')
scatterplot3d::scatterplot3d(x = data2$monocytes, y = data2$lymphocytes, z = data2$leukocytes,
                             pch = 16, color = colors[data2$admitted_to], angle = 55, type = 'h',
                             xlab = 'monocytes', ylab = 'lymphocytes', zlab = 'leukocytes')
legend(x = 'top', legend = levels(data2$admitted_to), col =  colors, pch = 16, 
       inset = -0.15, xpd = TRUE, horiz = TRUE)
```


## SECTION 2 DATA ANALYSIS METHODS 
### 2.1 TASK 1

## 2.1.1 Data Splitting

Splitting into training and testing
```{r}
set.seed(1)
trn_split_index <- createDataPartition(data_task1$task1_resp,p=0.7,
                                                 list = FALSE,times = 1)
task1_train<-data_task1[trn_split_index,]
task1_test<-data_task1[-trn_split_index,]
```

Checking data proportion in training set
```{r}
prop.table(table(task1_train$task1_resp))
```

## 2.1.2 MODELLING for imbalanced data

Defining fit control
```{r warning=FALSE,message=FALSE}
fit_ctrl <- trainControl(method = 'repeatedcv',
                     number = 10,
                     classProbs = TRUE,
                     summaryFunction = twoClassSummary)
```

### Model 2.1.2a - Logistic Regression 
```{r warning=FALSE,message=FALSE}
model_logist <- train(task1_resp~ .,task1_train,method = 'glm',
                      family = 'binomial',trControl = fit_ctrl,metric = 'ROC')
  
summary(model_logist)
print(model_logist)

predict_logist<-predict(model_logist,task1_test)

conf_mat_logist<-confusionMatrix(predict_logist,task1_test$task1_resp,dnn=c('Predicted','Observed'))
conf_mat_logist
```

Linear Discriminant Analysis (LDA) 
Figure 2.1.2b - Examining homogenity of variance by comparing covariate matrix ellipses for the +ve and-ve groups
```{r warning=FALSE,message=FALSE}
library(heplots)
task1_train %>% 
  dplyr::select(-c(1:4)) %>%
  heplots::covEllipses(., task1_train$task1_resp, pooled = F, labels = '', variables = 1:7, var.cex = 0.7)
```
It can be seen that we don't have the same covariance matrices for each class, so assumption is not satisifed.

Box's M-test for Covariance Matrices
```{r warning=FALSE,message=FALSE}
task1_train %>% 
  dplyr::select(-c(1:4)) %>%
  heplots::boxM(.,task1_train$task1_resp)
```
The Box’s M-test is rejecting the null hypothesis that the covariance matrices are equal, at 5% significant level.

### Model 2.1.2b - LDA Classifier
```{r warning=FALSE,message=FALSE}
model_lda <- train(task1_resp~.,task1_train[-3],
                   method= 'lda', trControl=fit_ctrl, metric='ROC')
model_lda

predict_lda<-predict(model_lda,task1_test)

confusionMatrix(predict_lda,task1_test$task1_resp,dnn=c('Predicted','Observed'))
```

Figure 2.1.2c - Variable Importance 
```{r}
imprt<-varImp(model_lda,scale = FALSE)
plot(imprt)
```

### Model 2.1.2c - QDA Classifier
```{r warning=FALSE,message=FALSE}
model_qda<-train(task1_resp~.,task1_train[,-3],method="qda",trControl=fit_ctrl,metric="ROC")
model_qda

predict_qda<-predict(model_qda,task1_test)
confusionMatrix(predict_qda,task1_test$task1_resp,dnn=c('Predicted','Observed'))
```

Figure 2.1.2d - Combining models
```{r}
comb_res <- resamples(list(Logistic=model_logist,LDA=model_lda,QDA=model_qda))

## plot
bwplot(comb_res)
```


### Model 2.1.2d - Random Forest Classifier
By default, number of trees is 500 and number of variables tried at each split is 3 in this case.
```{r warning=FALSE,message=FALSE}
model_rf<-train(task1_resp~.,task1_train,method='rf',trControl=fit_ctrl,metric="ROC")
print(model_rf)

pred_rf<-predict(model_rf,task1_test)
confusionMatrix(pred_rf,task1_test$task1_resp,dnn = c('Predicted','Observed'))
```

Figure 2.1.2e - Variable Importance 
```{r}
imprt<-varImp(model_rf,scale = FALSE)
plot(imprt)
```


## 2.1.3 DATA BALANCING using SMOTE

Defining fit control but with sampling method 
```{r warning=FALSE,message=FALSE}
fit_ctrl_smote <- trainControl(method = 'repeatedcv',
                     number = 10,
                     classProbs = TRUE,
                     summaryFunction = twoClassSummary,
                     sampling='smote')
```


## 2.1.4 Modeling for balanced data

### Model 2.1.4a - LOGISTIC with SMOTE
```{r warning=FALSE,message=FALSE}
model_logist_smt <- train(task1_resp~ .,task1_train,method = 'glm',
                      family = 'binomial',trControl = fit_ctrl_smote,metric = 'ROC')
  
model_logist_smt

predict_logist_smt<-predict(model_logist_smt,task1_test)

confusionMatrix(predict_logist_smt,task1_test$task1_resp,dnn = c('Predicted','Observed'))
```

### Model 2.1.4b - LDA with SMOTE
```{r warning=FALSE,message=FALSE}
model_lda_smt<-train(task1_resp~.,task1_train[-3],method="lda",trControl=fit_ctrl_smote,metric="ROC")
model_lda_smt

pred_lda_smt<-predict(model_lda_smt,task1_test,)
confusionMatrix(pred_lda_smt,task1_test$task1_resp,dnn=c('Predicted','Observed'))
```

### Model 2.1.4c - QDA with SMOTE
```{r warning=FALSE,message=FALSE}
model_qda_smt<-train(task1_resp~.,task1_train[-3],method="qda",trControl=fit_ctrl_smote,metric="ROC")
model_qda_smt

pred_qda_smt<-predict(model_qda_smt,task1_test)
confusionMatrix(pred_qda_smt,task1_test$task1_resp,dnn=c('Predicted','Observed'))
```

Combining results for classifiers with SMOTE using resamples
```{r}
comb_res_smote <- resamples(list(Logistic_SMOTE=model_logist_smt,LDA_SMOTE=model_lda_smt,
                           QDA_SMOTE=model_qda_smt))
```

### Model 2.1.4d - Random Forest Classifier with SMOTE
```{r warning=FALSE,message=FALSE}
model_rf_smt<-train(task1_resp~.,task1_train,method='rf',trControl=fit_ctrl_smote,,metric="ROC")
model_rf_smt

pred_rf_smt<-predict(model_rf_smt,task1_test)
confusionMatrix(pred_rf_smt,task1_test$task1_resp,dnn = c('Predicted','Observed'))
```

Figure 2.1.4a - Comparing classifiers with and without SMOTE
```{r include=TRUE,fig.height=6,fig.width=15}
orig_metr <- summary(comb_res)[[1]] %>% 
  pivot_longer(cols = 1:9) %>% 
  separate(col = 'name', into = c('model', 'metric'), sep = '~')

wid_smt_mtr <- summary(comb_res_smote)[[1]] %>% 
  pivot_longer(cols = 1:9) %>% 
  separate(col = 'name', into = c('model', 'metric'), sep = '~') %>% 
  mutate(model = paste0(model, ' (SMOTE)'))

orig_metr %>% 
  bind_rows(wid_smt_mtr) %>% 
  ggplot(aes(x = model, y = value, fill = model)) +
  stat_boxplot(geom = 'errorbar', size = 0.3) +
  geom_boxplot() + facet_wrap(~metric) +
  labs(title = 'Boxplot of metrics') + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())
```


### 2.2 TASK2

## 2.2.1 Data splitting for training and testing
```{r}
set.seed(1)
split_ind<-caret::createDataPartition(data2$admitted_to,p=0.7,list=F,times = 1)
task2_train<-data2[split_ind,]
task2_test<-data2[-split_ind,]
```

Checking proportions
```{r}
table(task2_train$admitted_to)
```


## 2.2.2 Model Construction

Defining fit control again as this time its a multi-classification problem
```{r warning=FALSE,message=FALSE}
fit_ctrl2 <- trainControl(method = 'repeatedcv',
                     number = 10,
                     classProbs = TRUE,
                     summaryFunction = multiClassSummary)
```


### Model 2.2.2a - MultiNOMIAL Model
```{r warning=FALSE,message=FALSE}
model_multinom<-train(admitted_to~.,task2_train,method='multinom',metric='AUC',
                      trControl=fit_ctrl2,trace=FALSE)
model_multinom

probs<-predict(model_multinom,task2_test,type='prob')
max_prob<-apply(probs,1,which.max)
pred_multinom<-factor(max_prob,labels=levels(task2_train$admitted_to))

confusionMatrix(pred_multinom,task2_test$admitted_to,dnn = c('Predicted','Observed'))
```
Figure 2.2.2a - Variable Importance for Multinomial Model 
```{r}
plot(varImp(model_multinom,scale=FALSE))
```


### Model 2.2.2b - Again with Random Forest classifier since it performed the best last time
```{r warning=FALSE,message=FALSE}
set.seed(123)
model_rf2<-train(admitted_to~.,task2_train,method='rf',metric='AUC',trControl=fit_ctrl2,trace=FALSE)
model_rf2

pred_rf2<-predict(model_rf2,task2_test)
confusionMatrix(pred_rf2,task2_test$admitted_to,dnn = c('Predicted','Observed'))
```


### Model 2.2.2c - SVM Non-Linear Classifier
```{r warning=FALSE,message=FALSE}
set.seed(123)
model_svm<-train(admitted_to~.,task2_train,method='svmRadial',
                 metric='AUC',trControl=fit_ctrl2)
model_svm

pred_svm<-predict(model_svm,task2_test)
confusionMatrix(pred_svm,task2_test$admitted_to,dnn = c('Predicted','Observed'))
```

```{r}
plot(model_svm)
```
At C=0.5, maximum AUC was obtained.

### Model 2.2.2d - KNN-Classifier
```{r warning=FALSE,message=FALSE}
set.seed(123)
model_knn<-train(admitted_to~.,task2_train,method='knn',metric='AUC',trControl=fit_ctrl2)
model_knn

pred_knn<-predict(model_knn,task2_test)
confusionMatrix(pred_knn,task2_test$admitted_to,dnn = c('Predicted','Observed'))
```

```{r}
plot(model_knn)
```

Figure 2.2.2b - Combining results
```{r fig.width=15}
comb_res_task2<-resamples(list(RForest=model_rf2,SVM=model_svm,Knn=model_knn))

#Figure 2.2.2a - Results Comparison plot
bwplot(comb_res_task2)
```

Figure 2.2.2c - Variable Importance for Best Classifier 
```{r}
plot(varImp(model_rf2,scale=FALSE))
```
In the feature importance plot, the top predcitors are the variables having highest distribution of response variable among them.


